---
title: "DATA602 Assignment2"
author: "An Yan  30167924"
date: "2022-09-27"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Question 1

```{r}
#part a
data1=read.csv('Data602Assignment1Question11.csv')
mean1=mean(data1$Delivery_time)
sd1=sd(data1$Delivery_time)
mean1
sd1
```

```{r}
pnorm(mean1, 5, (1.5)/(sqrt(12)))
```
The probability that another random sample of the same size will produce a sample mean that is at least the same value as the value of $\overline X$ observed in Question 11 of Assignment 1 is 0.9438.



```{r}
#part b
z1=11*(0.5^2)/sd1^2
z2=11*(1^2)/sd1^2
pchisq(z2, 11)-pchisq(z1, 11)
```
The probability that another random sample will yield a sample standard deviation that is between 0.5 hour and 1 hour is 0.0434.


# Question 2

```{r}
#part a
p2=0.42
n2=1426
n2*p2*(1-p2)
sd2=sqrt((p2*(1-p2))/(n2))
sd2
mean2=p2
mean2
```

The distribution shape is close to normal since np and np(1-p) are both grater than 10 by applying the central limit theorem, and the balancing point is equal to the proportion which is 0.42.

```{r}
#part b
pnorm(0.3794, p2, sd2)
```

```{r}
#part c
n2c = 1000
q2data = c(rep(0,(1-p2)*n2c), rep(1, p2*n2c))
q2boot = mosaic::do(2000) * mean(resample(q2data, n2c))
ggplot(data=q2boot, aes(x = mean)) + geom_histogram(col='blue', fill='red', binwidth=0.01) + xlab("Bootstrap Statistic - Sample Proportion")
```

# Question 3

```{r}
mean3=36/49
sd3=0.75998
pnorm(1, mean3, sd3/sqrt(52))
```

The random variable X has the hypergeometric distribution with $N=49\ M=6\  N=6$, by the CLT it will be normally distributed since it will be repeated 52 times. Thus I computed the probability that Billy claims that in 52 weeks on average he manages to have at least one matching number on his 6-49 ticket by compute 
$$
P(z>1)=P(\frac{1-(36/49)}{0.75998/\sqrt52})=0.9941
$$

# Question 4

```{r}
#part a
#create the bootstrap data
ntimes = 2000  
nsize = 12     
q4means = numeric(ntimes)  
q4sd = numeric(ntimes)   
origdata = c(16, 5, 21, 19, 10, 5, 8, 2, 7, 2, 4, 9) 
for(i in 1:ntimes) 
{   q4data = sample(origdata, nsize, replace=TRUE)  
    q4means[i] = mean(q4data)
        q4sd[i] = sd(q4data)       
} 
q4boot = data.frame(q4means)
```

```{r}
#part b
quantile(q4means, c(0.025, 0.975))
```

We can say that 95% of all possible unknown values of the population mean are between the 2.5-percentile and 97.5-percentile of the distribution of the bootstrap statistics  X_bar.

```{r}
#part c
t.test(q4means, conf.level=0.95)
```

Which means that 95% of the time the value of $\mu$ will fall between the confidence interval, also there is a 95% level of confidence that the unknown value of the population mean will be some point between the interval.

#part d

I will choose the second one, since t distribution is used when the sample size is small whereas is 12 in this question.

```{r}
#part e
#plot the bootstrap distribution of sample mean
ggplot(q4boot, aes(x = q4means)) + geom_histogram(col="red", fill="blue", binwidth=1) + xlab("Values of Bootstrap Statistic") + ylab("Count") + ggtitle("Histogram of Bootstrap Statistic: Sample mean")
```

Yes, the condition is satisfied. The bootstrap distribution in this histogram is very close to normal distribution although it is lightly right-skewed.

# Question 5

```{r}
#part a
X=571
n=1866
p=(X+2)/(n+4)
z5=qnorm(0.975)
lb5a=p-z5*sqrt((p*(1-p))/(n+4))
ub5a=p+z5*sqrt((p*(1-p))/(n+4))
cat('The 95% confidence interval for p is: [', lb5a, ',', ub5a, ']')
```

```{r}
#part b
#create the bootstrap data
nq5b = 2000
q5data = c(rep(0,(1-p)*nq5b), rep(1, p*nq5b))
q5boot = mosaic::do(2000) * mean(resample(q5data, nq5b))
ggplot(data=q5boot, aes(x = mean)) + geom_histogram(col='blue', fill='red', binwidth=0.01) + xlab("Bootstrap Statistic - Sample Proportion")
```

```{r}
#part c
qdata(~mean, c(0.025, 0.975), data=q5boot)
```

The 95% bootstrap confidence interval for p is [0.2865, 0.3265].

#part d 
The second one should be reported, even these two intervals are very close to each other, the bootstrap CI is based on a normally distributed sample.

# Question 6

```{r}
nhs=670
Xhs=348
nuni=376
Xuni=274
```

```{r}
#part a
phs=Xhs/nhs
#create the bootstrap data
nq6a = 1000
q6data = c(rep(0,(1-phs)*nq6a), rep(1, phs*nq6a))
q6boot = mosaic::do(1000) * mean(resample(q6data, nq6a))
ggplot(data=q6boot, aes(x = mean)) + geom_histogram(col='blue', fill='red', binwidth=0.01) + xlab("Bootstrap Statistic - Sample Proportion for High School or Less")
```

```{r}
#part b
puni=Xuni/nuni
#create the bootstrap data
n6b = 1000
q6bdata = c(rep(0,(1-puni)*n6b), rep(1, puni*n6b))
q6bboot = mosaic::do(1000) * mean(resample(q6bdata, n6b))
ggplot(data=q6bboot, aes(x = mean)) + geom_histogram(col='blue', fill='red', binwidth=0.01) + xlab("Bootstrap Statistic - Sample Proportion for University Degree")
```

```{r}
#part c
Nsims = 1000  
prop.uni = numeric(Nsims)
prop.hs = numeric(Nsims)
diff.props = numeric(Nsims)
for(i in 1:Nsims)
  {   prop.uni[i] = sum(sample(q6bdata, nuni, replace=TRUE))/nuni  
      prop.hs[i] = sum(sample(q6data, nhs, replace=TRUE))/nhs 
      diff.props[i] = prop.uni[i] - prop.hs[i] 
}
boot_prop = data.frame(prop.uni, prop.hs, diff.props)
```

```{r}
ggplot(data=boot_prop, aes(x = diff.props)) + geom_histogram(fill='blue', col='red', binwidth=0.025) + xlab("Distribution of Difference SamProp(uni) - SamProp(HS)") + ggtitle("Distribution of Bootstrap Stat: Sample Prop (University) - Sample Prop (HighSchool/Less)")
```

```{r}
qdata(~ diff.props, c(0.025, 0.975), data=boot_prop)
```

The 95% confidence interval of the difference between $P_{uni}$ and $P_{HS}$ is [0.1444, 0.2661].

```{r}
#part d
diff.props2 = numeric(Nsims)
for(i in 1:Nsims)
{ 
      diff.props2[i] = prop.hs[i] - prop.uni[i] 
}
boot_prop2 = data.frame(prop.uni, prop.hs, diff.props2)

ggplot(data=boot_prop2, aes(x = diff.props2)) + geom_histogram(fill='blue', col='red', binwidth=0.025) + xlab("Distribution of Difference SamProp(HS) - SamProp(uni)") + ggtitle("Distribution of Bootstrap Stat: Sample Prop (HighSchool/Less) - Sample Prop (University)")
```

```{r}
qdata(~ diff.props2, c(0.025, 0.975), data=boot_prop2)
```

```{r}
#plot a dot graph for two proportions
plot(prop.hs, prop.uni, type="p", col=c("blue", "red"))
```

From part c and d, the confidence interval for $P_{HS}-P{uni}$ is negative which means that the proportion of persons with at most high school education is greater than the proportion of persons with at least undergraduate university degree.

# Qustion 7

```{r}
#part a
#create the bootstrap data for median from Question 4
q4medians=numeric(ntimes)
for(i in 1:ntimes) 
{   q4data = sample(origdata, nsize, replace=TRUE)  
        q4medians[i]=median(q4data)
} 
q7boot_medians = data.frame(q4medians)
```

```{r}
quantile(q4medians, c(0.005, 0.995))
```

The 99% of the time the median will fall between the interval.

```{r}
#part b
q7boot_sds = data.frame(q4sd)
quantile(q4sd, c(0.025, 0.975))
```

# Question 8

```{r}
#part a
n=858
Xndp=352
pndp=Xndp/n
z8=qnorm(0.975)
lower=pndp-z8*(sqrt((pndp*(1-pndp))/(n)))
upper=pndp+z8*(sqrt((pndp*(1-pndp))/(n)))
cat('The 95% confidence interval for Pndp is: [', lower, ',', upper, '].')
```

```{r}
#part b
pndp2=(Xndp+2)/(n+4)
N8b = 1000
q8bdata = c(rep(0,(1-pndp2)*N8b), rep(1, pndp2*N8b))
q8bboot = mosaic::do(1000) * mean(resample(q8bdata, N8b))
ggplot(data=q8bboot, aes(x = mean)) + geom_histogram(col='blue', fill='red', binwidth=0.01) + xlab("Bootstrap Statistic for Pndp")
```

```{r}
#part c
lower2=pndp2-z8*(sqrt((pndp2*(1-pndp2))/(n)))
upper2=pndp2+z8*(sqrt((pndp2*(1-pndp2))/(n)))
cat("The 95% confidence interval for bootstrap statistic Pndp is: [", lower2, ',', upper2, '].')
```

#part d
We have 95% confidence to say that the proportion of Alberta voters who would vote for NDP will between 0.3778 and 0.4436 if a provincial election ere held "tomorrow".

