---
title: "DATA602 Assignment1"
output:
  pdf_document: default
  html_document:
    df_print: paged
date: "2022-09-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Question 1

```{r}
#part a
p=0.6
p_1a=p*p
p_1a
```

```{r}
#part b
p_1b=(1-p)^2
p_1b
```

```{r}
#part c
p1_c=2*p*(1-p)+p*p
p1_c
```

```{r}
#part d

```


##Question 2

```{r}
#Step 1
nsims=1000
outcome=numeric(nsims)
for(i in 1:nsims){ #we are going to perform the body of the loop 1000 times
  outcome[i] = sample(c(1,2,3,4,5,6), 1, replace=FALSE) #stores the outcome of the ith toss in position i of outcome
  }  #close the for-loop
simresult = data.frame(outcome) #creates a data frame with two columns
head(simresult,3)
tail(simresult,3)
```

```{r}
#Step 2
for(i in 1:nsims){ #we are going to perform the body of the loop 1000 times
  outcome[i] = sample(c(1,2,3,4,5,6), 1, replace=FALSE) #stores the outcome of the ith toss in position i of outcome
  }  #close the for-loop
simresult = data.frame(outcome) #creates a data frame with two columns
head(simresult,3)  #shows the first three rows of the data frame, outcome of each die toss in column 2
tail(simresult,3)  #shows the last three rows of the data frame, die toss outcome in column 2
```

```{r}
#Step 3

```

##Question 3
```{r}
#The sample space of this question is choosing 5 cards from a deck of 20 cards:
omega3=choose(20, 5)
#part a

#p=#A/#sample space
#number of A is 4 since there are four suits
p3a=4/omega3
p3a
```
The probability of straight flush is 0.0003.

```{r}
#part b
#compute the number of A: for example, start from heart:C(5,1), then 
A3b=choose(5,1)*choose(4, 3)*choose(4,2)*choose(4, 1)*choose(4,1)
p3b=A3b/omega3
p3b
```
The probability of three-of-a-kind is 0.1238.

```{r}
#part c
#
A3c=choose(3,1)*choose(4,1)*choose(12,2)+choose(3,2)*choose(4,2)*choose(12,1)
p3c=A3c/omega3
p3c
```
The probability that observes two Aces and two diamonds is 0.0650

##Question 4

Let AA denote the event flying Airline AA, UA denote the event flying Airline UA, D denote the event flying Airline D; and M denotes the event missing connecting flight.
In this question, we are computing the conditional probability:
$p(flying\ Airline\ UA| missing\ connecting\ flights)=p(UA|M)=\frac {p(UA\cap M)} {p(M)}$
thus we need to find $p(M\cap A)$ and $p(UA)$:
```{r}
#First compute the probabilities of event that flying three airlines respectively:
paa=0.15
pua=0.85/4
pd=3*pua
pua
pd
```
Thus $p(AA)=0.15$, $p(UA)=0.2125$ and $p(D)=0.6375$.
And we know $p(M|AA)=0.15$, $p=(M|UA)=0.3$ and $p=(M|D)=0.1$, we can use total probability theorem to calculate $p(M): p(M)=\sum_{n}p(M|AIRLINE_i)p(AIRLINE_i)$
```{r}
pm=0.15*0.15+0.2125*0.3+0.6375*0.1
pm
```
Now we know $p(M)=0.235$, and $p(M\cap UA)=p(UA)*p(M|UA)$:
```{r}
pmua=0.3*0.2125
pmua
```
Finally we can compute the conditional probability: $p(UA|M)=\frac {p(UA\cap) M}{p(M)}$
```{r}
puam=pmua/pm
puam
```
The probability that the executive called from Chicago given that the group has missed the connecting flight is 0.425.


##Question 5
```{r}
#part a
den5=function(x)(2/(3^(x+1)))
xvalues=0:15
```

```{r}
# part b
integrate(den5, lower=3, upper=15)
```

```{r}
#part c
mean5=integrate(function(x)(2*x/(3^(x+1))), lower=0, upper=15)
mean5=mean5$value
mean5
```

```{r}
# part d
square5=integrate(function(x)((2*x^2)/(3^(x+1))), lower=0, upper=15)
square5=(square5$value)
sd5=square5-mean5^2
sd5
```

```{r}
#part e
p5e=integrate(function(x)(2*x/(3^(x+1))), lower=mean5-sd5, upper=mean5+sd5)
p5e
```

##Question 8

```{r}
mean8=1.89
sd8=0.05
```

We are computing the probability: $P(1.83<x<1.91)$
```{r}
pnorm(1.91, mean8, sd8)-pnorm(1.83, mean8, sd8)
```
The probability that the amount of soft-drink dispensed into this bottle is between 1.83 and 1.91 litres is 0.5404.

```{r}
#part b
quantile(dnorm(mean8, sd8),0.9)
```
The 90th-percentile is 0.0734

```{r}
#part c
1-pnorm(2, mean8, sd8)
```
The proportion of all 2-litre bottles will be filled to overflow is 0.0139.

First we need to compute the probability that bottles will have less than 1.85 L of soft drink: $p(x<1.85)$
```{r}
#part d
p8d=pnorm(1.85, mean8, sd8)
```
Since we picked 50 bottles randomly, it's a binomial distribution: $Y\sim Binomal(50,\ p8d)$, we can use it to calculate $P(5\leq x \leq 10)$
```{r}
pbinom(10, size=50, prob=p8d)-pbinom(5, size=50, prob=p8d)
```
The probability that between 5 and 10 of these bottles will have less than 1.85 litres of soft-drink is 0.4691.


##Question 9
```{r}
gss=read.csv("http://people.ucalgary.ca/~jbstall/DataFiles/GSS2002.csv")
head(gss, 5)
```

##Question 10

```{r}
library(ISLR)
head(Default, 4)
```

```{r}
#part a
ggplot(data=Default, aes(x = balance, y = income, color=student)) + geom_point(size=2, position="jitter") + xlab("Balance") + ylab("Income") + ggtitle("Scatterplot of Annual Income to Credit Card Balancing")
```

```{r}
#part b
ggplot(data=filter(Default, student =="Yes" | student=="No" ), aes(x = student, y = balance)) + geom_boxplot(color = "orange", fill="blue", na.rm=TRUE) + coord_flip() + xlab("Balance Owing") + ylab("Student") + ggtitle("Boxplots of Balance Owing")
```

```{r}
#part c
#mean
mean10=mean(Default$balance)
mean10
#median
median10=median(Default$balance)
median10
#sd
sd10=sqrt(var(Default$balance))
sd10
#5th and 95th percentiles
quantiles10=quantile(Default$balance, c(0.05, 0.95))
quantiles10
```

The sample mean is 835.3749, the sample median is 823.637, the sample standard deviation is 483.715, the 5th and 95th percentiles are 0.4247 and 1665.96258 respectively.

##Question 11
```{r}
data11=read.csv('Data602Assignment1Question11.csv')
head(data11, 5)
```


```{r}
#part a
ggplot(data11, aes(x = Delivery_time))
```

```{r}
#part b
#sample mean
mean11=mean(data11$Delivery_time)
mean11
#sample median
median11=median(data11$Delivery_time)
median11
# sample sd
var11=var(data11$Delivery_time)
sd11=sqrt(var11)
sd11
#percentiles
quantile11=quantile(data11$Delivery_time, c(0.25, 0.75, 0.99))
```
The sample mean is 5.6875, the sample median is 5.775, the sample standard deviation is 1.5804, the first,  third and 99th percentiles are 4.75, 6.565 and 7.9778 respectively.

```{r}
#part c

```




